### 2024-09-19 cognitive-bias.qmd
20240918 - loss aversion and endowment effect

*Loss aversion bias* is the tendency to avoid losses rather than acquiring equivalent gains.
That is, loss aversion (similar to punishment sensitivity, as opposed to reward sensitivity) is the tendency to weigh losses (or punishment) more heavily than gains (or rewards).
Loss aversion is a key aspect of prospect theory [@Kahneman2011].
The loss aversion ratio indicates the extent to which a person weighs losses greater than gains, and for many people is between 1.5 to 2.5 [@Kahneman2011].
Loss aversion is different from [risk aversion](#sec-cognitiveBiasesRiskAversion).
Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
In fantasy football, loss aversion may lead managers to start or hold onto players for too long who were highly drafted yet are underperforming instead of starting a more promising player out of fear of losing potential value from their initial investment.

@@

*Endowment effect bias* is the tendency to overvalue merely because one owns it.
That is, a manager tends to value the same player more when they play for their team rather than an opponent's team.
When trading, the endowment effect bias might lead a manager to demand more to trade away a player than to acquire the same player [@Miller2013c].
For instance, a manager might overvalue a player they drafted in the first round, refusing to trade them even if they could get a better-performing player in return.

>>>>

*Loss aversion bias* is the tendency to avoid losses rather than acquiring equivalent gains.
That is, loss aversion (similar to punishment sensitivity, as opposed to reward sensitivity) is the tendency to weigh losses (or punishment) more heavily than gains (or rewards).
Thus, the target state matters less than the *change* (i.e., gains and/or losses) from one's reference point [e.g., their current roster; @Kahneman2011].
Loss aversion is a key aspect of prospect theory and suggests that individuals tend to be biased in favor of their reference situation and toward making smaller rather than larger changes [@Kahneman2011].
Loss aversion explains why people tend to favor the favoring the status quo; the negative consequences of a change often loom larger than the positive consequences of a change [@Kahneman2011].

The loss aversion ratio indicates the extent to which a person weighs losses greater than gains, and for many people is between 1.5 to 2.5 [@Kahneman2011].
Loss aversion is different from [risk aversion](#sec-cognitiveBiasesRiskAversion).
Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
In fantasy football, loss aversion may lead managers to start or hold onto players for too long who were highly drafted yet are underperforming instead of starting a more promising player out of fear of losing potential value from their initial investment.

@@

*Endowment effect bias* is the tendency to overvalue merely because one owns it—owning it increases its value to the owner [especially for goods that are not regularly traded; @Kahneman2011].
That is, a manager tends to value the same player more when they play for their team rather than an opponent's team.
It is related to [loss aversion bias](#sec-cognitiveBiasesLossAversion) and prospect theory; when owning a good (i.e., a person's reference point), they focus on the *loss* of the good if they were they to sell or trade it away rather than on the gains from the sale or trade.
When trading, the endowment effect bias might lead a manager to demand more to trade away a player than to acquire the same player [@Miller2013c].
For instance, a manager might overvalue a player they drafted in the first round, refusing to trade them even if they could get a better-performing player in return.
